### 2016-05-05 doc/2016-05 Luchon /presentation_SIG_Toulouse_Avril_2016.tex

\documentclass{beamer}

\usepackage{default}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}

\DeclareGraphicsExtensions{{.pdf},{.png},{.jpg}}
\graphicspath{ {./img/} }

\hypersetup{colorlinks,urlcolor=blue}
\AtBeginSection[]
{
	\begin{frame}
		\begin{NoHyper}
		\tableofcontents[currentsection]
		\end{NoHyper}
	\end{frame}
}
\AtBeginSubsection[]
{
	\begin{frame}
		\begin{NoHyper}
		\tableofcontents[currentsection, currentsubsection]
		\end{NoHyper}
	\end{frame}
}
\AtBeginSubsubsection[]
{
	\begin{frame}
		\begin{NoHyper}
		\tableofcontents[currentsection, currentsusubbsection]
		\end{NoHyper}
	\end{frame}
}
\begin{document}

\addtobeamertemplate{footline}{\insertframenumber/\inserttotalframenumber}

\title{You are the way you (structurally) talk:  Structural-temporal neighbourhoods of posts to characterize users in online forums}
\author{Alberto Lumbreras \\Jouve B., Velcin J., Guégan, M.}
\date{May 3, 2016}
\maketitle

\begin{frame}\frametitle{Overview} 
\begin{NoHyper}
\tableofcontents
\end{NoHyper}
\end{frame}

\section{Introduction}
\subsection{The data}

\begin{frame}{The data}{Reddit. A forum of forums}
	\begin{figure}
		\centering
		\includegraphics[width=0.1\textwidth]{reddit-logo}	
	\end{figure}
	Download monthly dumps from:
	\href{http://couch.whatbox.ca:36975/reddit/comments/monthly/}{http://couch.whatbox.ca:36975/reddit/comments/monthly/}

	\vfill
	Extract forum of interest:\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}science}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}france}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}sociology}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}complexsystems}}\\
	\href{www.reddit.com/r/gameofthrones}{www.reddit.com/r/{\color{red}gameofthrones}} $\leftarrow$ in this presentation\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}podemos}} $\leftarrow$ in this presentation\\
	...
\end{frame}

\subsection{The graph representations of the data}

\begin{frame}{Graph representations}{Graph of user interactions (a social network)}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{sna}	
	\end{figure}
\end{frame}

\begin{frame}{Graph representations}{Trees of posts}
\begin{figure}
	\centering
	\includegraphics[width=1\textwidth]{forest}	
\end{figure}
\end{frame}

\begin{frame}{Graph representations}{Conversations are trees}

\begin{itemize}
\item Explicit structure.
\item Dynamic (order, time)
\end{itemize}
\begin{figure}
	\centering
	\includegraphics[width=0.3\textwidth]{tree1}	
	\includegraphics[width=0.3\textwidth]{tree2}	
	\includegraphics[width=0.3\textwidth]{tree4}	
\end{figure}
\end{frame}

\section{Structures of conversations}
\subsection{Basic idea}
\begin{frame}{Intuition}
\textit{Hypothesis}: different individuals have tendency towards different types of conversations and these types are reflected in the structure of their interactions.

\end{frame}

\subsection{Triadic structures}
\begin{frame}{Triadic structures}{Triads are not enough}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{triads}
	\end{figure}

Triads in \textbf{trees of posts}:
	\begin{itemize}
		\item Only 3 possible triads (dyad, chain and star)
	\end{itemize}

Triads in  \textbf{social graph}:
	\begin{itemize}
		\item Order (therefore dynamic) is missing.
	\end{itemize}
\vfill
We need something richer that captures the dynamics of conversations.
\end{frame}

\subsection{Neighbourhood structures}
\begin{frame}{Order-based neighbourhoods}{Definition}
\begin{itemize}
	\item 1. Extract neighbourhood of post $i$ with radius $r$.
	\item 2. Keep only the $n$ posts that are closest (in time) to post $i$. 
\end{itemize}
	\begin{figure}
		\centering
		\includegraphics[width=01\textwidth]{order_neighbourhood}
	\end{figure}
Easy to compute, but how to choose the parameters $r,n$ ?  	
\end{frame}

\begin{frame}{Time-based neighbourhoods}{Definition}
\begin{itemize}
	\item 1. Extract neighbourhood of post $i$ with radius $r$.
	\item 2. Detect changes of speed (vertical/horizontal \textit{changepoints}) (PELT algo)
	\item 3. From $i$, get the posts around until a changepoint is found.
\end{itemize}
%\vspace{cm}  
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{breakpoints2}
	\end{figure}
\end{frame}

\begin{frame}{Colouring}
Label special nodes:
\begin{itemize}
	\item Red: ego.
	\item Yellow: parent of ego (and posts of same author)
	\item Orange: other posts by ego author
	\item White: root
\end{itemize}
	\begin{figure}
		\centering
		\includegraphics[width=0.3\textwidth]{colouring}
	\end{figure}

\end{frame}
\begin{frame}{Some frequent neighbourhoods}{}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{dictionary1}
	\end{figure}
\end{frame}
\begin{frame}{Some frequent neighbourhoods}{}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{dictionary2}
	\end{figure}
\end{frame}

\subsection{Comparing neighbourhoods}
\begin{frame}{Frequency distribution}{}
	\begin{figure}
		\centering
		\includegraphics[width=0.8\textwidth]{compare_census_distribution}
	\end{figure}
Time-based (black) reduces the space w.r.t structure-based (blue)
\end{frame}

\begin{frame}{Frequency distribution}{}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{compare_census_distribution_cum}
	\end{figure}
\end{frame}

\begin{frame}{Discrepancies}{Time-based vs Order-based neighbourhood}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{confusion_time_order}
	\end{figure}
\end{frame}

\begin{frame}{Discrepancies}{Basic neighbourhood vs Time-based neighbourhood}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{confusion_struct_time}
	\end{figure}
\end{frame}

\begin{frame}{Structure-based vs Order-based vs Time-based}
Structure-based:
\begin{itemize}
\item  too big (and too many) neighbourhoods.
\end{itemize}
Order-based:
\begin{itemize}
\item  Dominance of monoid hides richer conversational structures.
\end{itemize}
Time-based:
\begin{itemize}
\item Space more reduced than simple structure-based.
\item Criteria to choose radius dynamically (r = until conversation slows down)
\end{itemize}
\end{frame}

\section{Conversation-based clustering}
\begin{frame}{Methodology}
	\begin{itemize}
	\item Create a user $\times$ neighborhood matrix of counts.
	\item Z-normalize (users characterized by their deviation from the mean)
	\item Cluster!
\end{itemize}
\end{frame}

\begin{frame}{Conversation-based clustering}{Time-based}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{cluster_whiskers}
	\end{figure}	
\end{frame}

\begin{frame}{Conversation-based clustering}{Interpretation}
	Greens:
	\begin{figure}
		\centering
		\includegraphics[height=0.17\textwidth]{motif8}
		\includegraphics[height=0.17\textwidth]{motif4}
		\includegraphics[height=0.17\textwidth]{motif15}
	\end{figure}	
	Reds:
	\begin{figure}
		\centering
		\includegraphics[height=0.17\textwidth]{motif20}
		\includegraphics[height=0.17\textwidth]{motif3}
		\includegraphics[height=0.17\textwidth]{motif19}	
		\includegraphics[height=0.17\textwidth]{motif14}	
		\includegraphics[height=0.17\textwidth]{motif13}	
		\includegraphics[height=0.17\textwidth]{motif18}	
	\end{figure}	
	Blues (avoid these motifs):
	\begin{figure}
		\centering
		\includegraphics[height=0.17\textwidth]{motif3}
		\includegraphics[height=0.17\textwidth]{motif19}
	\end{figure}	
	\end{frame}

\section{Conclusions}
\begin{frame}{Conclusions}
	\begin{itemize}
	\item \textbf{Q: Can we use graph structure to characterise users?}
	\item A: Yes!
	\end{itemize}	
	\begin{itemize}
	\item \textbf{Q: By using triads}?
	\item A: No. They are not useful in trees.
	\end{itemize}
	\begin{itemize}
	\item \textbf{Q: So, what kind of structure?}
	\item A: Posts neighbourhoods that are time/order sensitive.
	\end{itemize}
	\begin{itemize}
	\item \textbf{Q: What about language?}
	\item A: It's ok, but structure is more directly linked to thread dynamics (future work)
	\end{itemize}
\end{frame}

\begin{frame}
\begin{center}
\Huge Merci !
	\begin{figure}
		\centering
		%\includegraphics[width=0.1\textwidth]{reddit-logo}
		%\includegraphics[width=0.2\textwidth]{reddit_french}	
		%\includegraphics[width=0.5\textwidth]{reddit_questions}
		\includegraphics[width=0.1\textwidth]{reddit_sans_culotte}
	\end{figure}
\end{center}
\end{frame}
\end{document}